\chapter{Methodology}\label{C:method} 
The core of this empirical study is the analysis of corpora of code written in each of the investigated languages. This analysis makes use of many static code analysis methods including the following:
\begin{itemize}
	\item Grep is used to perform regular expression searches on files. This can detect the more simple patterns explored in this paper.
	\item ANTLR is a tool which accepts a language grammar and a valid file in that language. From this it constructs a syntax tree to represent the file.
	\item JSClassFinder is a tool which detects class and method declarations in JavaScript code.
	\item Esprima accepts a JavaScript file as input and produces a JSON representation of the syntax tree of that file. This JSON file is then used as the input to JSClassFinder.
\end{itemize}
Each of these tools helps to extract valuable information from one of more of the languages analysed in this study.

\section{Selecting Languages}
The languages which have been explored thus far are limited to Java and JavaScript. These languages were covered first because they both have large open source communities and they differ on their native method of object inheritance which provides a good avenue for comparison. The other languages which will be explored in the remainder of the project are Python, Lua and Scala. These are included to ensure that a wide enough variety of languages are analysed that it is possible to make conclusions about the usage of delegation and inheritance across languages.

\section{Assembling Corpora}
To analyse each language, we first needed to collect a corpus representative of that language's use in real world software development projects. In the case of Java, we adopted The Qualitas Corpus, which is a large collection of open source projects written in the Java language~\cite{QualitasCorpus}. Likewise, with JavaScript, we have adopted an existing corpus used by the team that developed JSClassFinder~\cite{JSClassFinder}.
\newline

For the other studied languages, Python, Lua, and Scala, quality existing corpora could not be found. For each of these languages, the top 25 open source projects were sourced from GitHub's "Trending this month" list for June, 2016. This source was chosen because it provides a group of projects for each language which are in active development as measured by GitHub, and which are easy to access. This helps to ensure that the analysis performed will be as relevant as possible to modern software development.

\section{C\# Analysis}
The ANTLR analysis for a project works in three major passes:
\begin{enumerate}
	\item Create a syntax tree for each code file and traverse it to find all class declaration subtrees. Collect these class declarations to use in later steps.
	\item Run a visitor down each class declaration subtree, searching for all the methods and recording their modifiers. A type hierarchy is also established at this step to allow classes to find information about method calls they make which may be dispatched to a method in their superclass.
	\item Run another visitor down each class declaration tree and find constructors and check which methods are called against the modifiers found in the previous pass to determine which methods could miss their intended target under a different object initialisation model.
\end{enumerate}






